### 2024-03-26 thesis/input_files/probabilistic.tex
March 26th 2024

        \frac{\partial L}{\partial b}=0 \implies \alpha-\alpha^*=0

@@

Where an unbiased estimator for $b$ is given by

@@

    b=&\frac{1}{I_q}\sum\limits_{i\in I_q} b_i, \\ 
    \textrm{with} \ b_i=& y_i-\sum\limits_k \alpha_k k(x_k, x_i) \\
     \textrm{and} \ I_q=& \{i=1,\dots,n|0<\alpha_i<Cq, 0<\alpha_i<C(1-q)\}

@@

%contribution is implementation kernel quantile regression for pythos users since there exists only r version
\subsection{Weather quantiles}
In order to get acquainted with the inner workings of the presented methods, this section covers an application explaining practical details and comparing results.

@@

The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}. It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
Due to the bimodality of the data, such dataset is commonly used to give a difficult quantile regression problem \cite{hyndman1996estimating}, thus why we chose it. The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
\begin{figure}
    \includegraphics[width=\textwidth]{images/melbourne_temperature.png}
    \caption{Melbourne temperatures dataset}
    \label{fig:melbourne_temperature_data}
\end{figure}
Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation. Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
\begin{figure}
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_linear_quantile_regression.png} 
      \caption{Linear quantile regression} 
      \label{fig:melbourne_linear_quantile_regression} 
      \vspace{4ex}
    \end{subfigure}%% 
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_gradient_boosting_quantile_regression.png} 
      \caption{Gradient boosting quantile regression} 
      \label{fig:melbourne_gradient_boosting_quantile_regression} 
      \vspace{4ex}
    \end{subfigure} 
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_quantile_forest.png} 
      \caption{Quantile forest} 
      \label{fig:melbourne_quantile_forest} 
    \end{subfigure}%%
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melborune_kernel_quantile_regression.png} 
      \caption{Kernel quantile regression} 
      \label{fig:melborune_kernel_quantile_regression} 
    \end{subfigure} 
    \caption{Quantile regressors}
    \label{fig:melbourne_quantiles_comparison} 
  \end{figure}

>>>>

        \frac{\partial L}{\partial b}=0 \implies (\alpha-\alpha^*)\mathbb{1}=0

@@

Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions, see section 5.5.3 \cite{boyd2004convex}.
Thanks the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
To see this, we have to consider the KKT conditions.

@@

    \lambda^* r_i(x^*)=&0, \quad i=1,\dots,m
    \nabla L(x^*)=&0

@@

In our setting we have
        \alpha_i(\xi_i-y_i+r_i)=&0

@@

        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
        \nu_i \xi_i=&0
        \nu_i^* \xi_i^*=&0
        \nabla L =&0
Using the gradient of the lagrangian of equation \ref{eq:lagrange_derivatives} we end up with
        \alpha_i(\xi_i-y_i+r_i)=&0
        \alpha_i^*(\xi_i^*+y_i-r_i)=&0
        (Cq-\alpha_i) \xi_i=&0
        (C(1-q)-\alpha_i) \xi_i^*=&0

Now, let us break into cases

\begin{equation}
\begin{cases}
    \alpha_i=Cq, \ \alpha_i^*=0 & \implies \alpha_i-\alpha_i^*=Cq, \xi_i\leq 0, \xi^*=0\implies \xi_i-y_i+f_i=0 \\
    \alpha_i=0, \ \alpha_i^*=C(1-q) & \implies \alpha_i-\alpha_i^*=C(q-1), \xi_i= 0, \xi^*\leq 0 \implies \xi_i^*+y_i-f_i=0 \\
    0\leq \alpha_i\leq Cq, \ 0\leq \alpha_i^*\leq C(1-q) & \implies \xi_i=0,  \xi^*_i= 0 \implies -y_i+f_i=0 \ y_i-f_i=0 \\
\end{cases}
\end{equation}

Therefore in order to retrieve b we simply have to choose an index $i$ such that bold $\alpha_i \not \in \{C(q-1), Cq\}$
and let
\begin{equation}
    \begin{aligned}
    b=&y_i-\sum_i \alpha_i k(x_i,x)
    \end{aligned}
\end{equation}

% In our implementatiton we take the point that minsimizes the distance between the boundary of domain of alpha, in this way we are guaranteed we do not take ith index of alpha_i for retrieving b, because it that case it would not work.
%contribution is implementation kernel quantile regression for pythos users since there exists only r version
\subsection{Weather quantiles}
In order to get acquainted with the inner workings of the presented methods, this section covers an application explaining practical details and comparing results.
\\
The dataset used is the Melbourne Daily Maximum Temperatures \cite{hdrcde}. It contains the daily maximum temperatures in Melbourne, Australia, from 1981-1990, excluding leap days, see figure \ref{fig:melbourne_temperature_data}.
Due to the bimodality of the data, such dataset is commonly used to give a difficult quantile regression problem \cite{hyndman1996estimating}, thus why we chose it. The observed bimodality is that a hot day is likely to be followed by either an equally hot day or one much cooler.
\begin{figure}
    \includegraphics[width=\textwidth]{images/melbourne_temperature.png}
    \caption{Melbourne temperatures dataset}
    \label{fig:melbourne_temperature_data}
\end{figure}
Hereafter, the results of the four presented methods on the Melbourne dataset are reported, see figure \ref{fig:melbourne_quantiles_comparison} for a visualisation. Hyperparameter tuning have been optimised through cross validation, see \ref{appendix:cross_validation}.
\begin{figure}
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_linear_quantile_regression.png} 
      \caption{Linear quantile regression} 
      \label{fig:melbourne_linear_quantile_regression} 
      \vspace{4ex}
    \end{subfigure}%% 
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_gradient_boosting_quantile_regression.png} 
      \caption{Gradient boosting quantile regression} 
      \label{fig:melbourne_gradient_boosting_quantile_regression} 
      \vspace{4ex}
    \end{subfigure} 
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melbourne_quantile_forest.png} 
      \caption{Quantile forest} 
      \label{fig:melbourne_quantile_forest} 
    \end{subfigure}%%
    \begin{subfigure}[b]{0.5\linewidth}
      \centering
      \includegraphics[width=1.1\textwidth]{images/melborune_kernel_quantile_regression.png} 
      \caption{Kernel quantile regression} 
      \label{fig:melborune_kernel_quantile_regression} 
    \end{subfigure} 
    \caption{Quantile regressors}
    \label{fig:melbourne_quantiles_comparison} 
  \end{figure}
